Wi-Fi Based Handwritten Signature Verification Using a Triplet Network

### 2019-10-01 190906_Paper_BDML_v1/bdml_v0.tex
paper_bdml: v2

% Karnet structure and mining samples.
Given $m$ training samples. Let $\mathbf{X}\in{\mathbb{R}}^{m \times (n+1)}$ denotes the training data set and $\mathbf{G}\in{\mathbb{R}}^{m \times n}$ denotes the network output.
Then the multilayer neural network can be written in equation form as follows:
\begin{equation}
    \mathbf{G}=\sigma\left(\left[\mathbf{1},\sigma\left(\dots\left[\mathbf{1},\sigma\left(\left[\mathbf{1},\sigma\left(\mathbf{X}\mathbf{W}_{1}\right)\right]\mathbf{W}_{2}\right)\right]\dots\mathbf{W}_{(n-1)}\right)\right]\mathbf{W}_{n}\right),
\end{equation}
where $\mathbf{W}_{1}\in{\mathbb{R}}^{(n+1) \times h_{1}}$,$\mathbf{W}_{2}\in{\mathbb{R}}^{(h_{1}+1) \times h_{2}}$,$\dots,\mathbf{W}_{n}\in{\mathbb{R}}^{(h_{(n-1)}+1) \times n}$ are the network weight matrices, $\mathbf{1}=\left[1,\dots,1\right]^{T}\in{\mathbb{R}}^{m \times 1}$ is the bias vector, and $\sigma(.)$ is the activation function.
By adopting an one-hot encoded target $\mathbf{Y}\in{\mathbb{R}}^{m \times n}$, training of the weight matrices $\mathbf{W}_{i}$ using the KAR space method~\cite{toh2018gradient} can be computed as follows:
\begin{IEEEeqnarray}{rCl}
    \mathbf{W}_{i}&=&\left[\mathbf{1},\sigma\left(\dots\left[\mathbf{1},\sigma\left(\left[\mathbf{1},\sigma\left(\mathbf{X}\mathbf{W}_{1}\right)\right]\mathbf{W}_{2}\right)\right]\dots\mathbf{W}_{(i-1)}\right)\right]^{\dagger}\sigma^{-1}\left(\mathbf{Y}\right), \nonumber \\ i&=&1,\dotsc,n.
\end{IEEEeqnarray}

@@

By training the network with the single shot KAR space learning, we can map the $L2$ distance between every sample by using the output vector of the KAR space network. Given a set of training data which is packed in matrix $\mathbf{X}$, the network output can be written as:
\begin{equation}
    f\left(\mathbf{X}\right)=\sigma\left(\left[\mathbf{1},\sigma\left(\dots\left[\mathbf{1},\sigma\left(\left[\mathbf{1},\sigma\left(\mathbf{X}\cdot\mathbf{W}_{1}\right)\right]\mathbf{W}_{2}\right)\right]\dots\mathbf{W}_{(n-1)}\right)\right]\mathbf{W}_{n}\right).
\end{equation}
After training based on the KAR space projection, the given anchor sample $\mathbf{x}_{anc}$ can be used as the reference to decide whether the network output of a sample is far away from this anchor sample. In other words, a hard positive sample $\mathbf{x}_{pos}$ and a hard negative sample $\mathbf{x}_{neg}$ with respect to the anchor sample $\mathbf{x}_{anc}$ can each be determined based on:
\begin{equation}
    {\left\| {{f\left(\mathbf{x}_{anc}\right)} - {f\left(\mathbf{x}_{pos}\right)}} \right\|_2^2} \geq \mathrm{t}_{pos}, \label{pos}
\end{equation}
\begin{equation}
    {\left\| {{f\left(\mathbf{x}_{anc}\right)} - {f\left(\mathbf{x}_{neg}\right)}} \right\|_2^2} \leq \mathrm{t}_{neg},\label{neg}
\end{equation}
where $\mathrm{t}_{pos}$ and $\mathrm{t}_{neg}$ respectively denote the thresholds that determine whether a sample is hard positive or hard negative. Since the hardest samples are likely to be outliers which can degrade the training process of the triplet network, we empirically set the $\mathrm{t}_{pos}$ at 75 percentile of the $L2$ distance and $\mathrm{t}_{neg}$ at 25 percentile of the $L2$ distance. The final set of $\mathbf{x}_{pos}$ and $\mathbf{x}_{neg}$ samples is randomly selected based on equations~\eqref{pos} and~\eqref{neg}.

@@

\subsubsection{Performance evaluation:}
The proposed system is evaluated under two cases: i) case I on comparison between the proposed system and other handcraft or deep learning-based methods based on the verification accuracy, and case II on detailed comparison between the proposed system and the deep learning-based methods using the receiver operating characteristic (ROC) curve and training loss curve. For both cases, Wi-Fi signal measurements based on a single-orientation and 4-orientations of the user are recorded. For Case I, existing handcraft methods such as the least squares error estimation (LSE)~\cite{duda2012pattern}, the principal components analysis (PCA)~\cite{turk1991eigenfaces} with LSE, the support vector machine (SVM)~\cite{vapnik2013nature} with different kernel functions, the total error rate minimization which adopted the reduced multivariate polynomial model as basis function (TER-RM2) \cite{toh2003fingerprint,toh2008between}, the deep learning-based Siamese network \cite{koch2015siamese}, and the baseline triplet network \cite{hoffer2015deep} are included for performance benchmarking. The Siamese network and the baseline triplet network utilized the same ConvNet structures with the proposed system and only differed in their input data style and loss function.

>>>>

% Karnet structure and mining samples.
Given $m$ training samples. Let $\mathbf{X}\in{\mathbb{R}}^{m \times (n+1)}$ denotes the training data set and $\mathbf{G}\in{\mathbb{R}}^{m \times n}$ denotes the network output.
Then the multilayer neural network can be written in equation form as follows:
\begin{IEEEeqnarray}{rCl}
    \mathbf{G}&=&\sigma\left(\left[\mathbf{1},\sigma\left(\dots\left[\mathbf{1},\sigma\left(\left[\mathbf{1},\sigma\left(\mathbf{X}\mathbf{W}_{1}\right)\right]\mathbf{W}_{2}\right)\right]\dots\mathbf{W}_{(i-1)}\right)\right]\mathbf{W}_{i}\right), \nonumber \\ i&=&1,\dotsc,n,
\end{IEEEeqnarray}
where $\mathbf{W}_{1}\in{\mathbb{R}}^{(n+1) \times h_{1}}$,$\mathbf{W}_{2}\in{\mathbb{R}}^{(h_{1}+1) \times h_{2}}$,$\dots,\mathbf{W}_{i}\in{\mathbb{R}}^{(h_{(i-1)}+1) \times n}$ are the network weight matrices, $\mathbf{1}=\left[1,\dots,1\right]^{T}\in{\mathbb{R}}^{m \times 1}$ is the bias vector, and $\sigma(.)$ is the activation function.
By adopting an one-hot encoded target $\mathbf{Y}\in{\mathbb{R}}^{m \times n}$, training of the weight matrices $\mathbf{W}_{i}$ using the KAR space method~\cite{toh2018gradient} can be computed as follows:
\begin{IEEEeqnarray}{rCl}
    \mathbf{W}_{i}&=&\left[\mathbf{1},\sigma\left(\dots\left[\mathbf{1},\sigma\left(\left[\mathbf{1},\sigma\left(\mathbf{X}\mathbf{W}_{1}\right)\right]\mathbf{W}_{2}\right)\right]\dots\mathbf{W}_{(i-1)}\right)\right]^{\dagger}\sigma^{-1}\left(\mathbf{Y}\right), \nonumber \\ i&=&1,\dotsc,n.
\end{IEEEeqnarray}

@@

By training the network with the single shot KAR space learning, we can map the $L2$ distance between every sample by using the output vector of the KAR space network. Given a set of training data which is packed in matrix $\mathbf{X}$, the network output can be written as:
\begin{IEEEeqnarray}{rCl}
    f\left(\mathbf{X}\right) &=& \sigma\left(\left[\mathbf{1},\sigma\left(\dots\left[\mathbf{1},\sigma\left(\left[\mathbf{1},\sigma\left(\mathbf{X}\cdot\mathbf{W}_{1}\right)\right]\mathbf{W}_{2}\right)\right]\dots\mathbf{W}_{(i-1)}\right)\right]\mathbf{W}_{i}\right), \nonumber \\ i&=&1,\dotsc,n.
\end{IEEEeqnarray}
After training based on the KAR space projection, the given anchor sample $\mathbf{x}_{anc}$ can be used as the reference to decide whether the network output of a sample is far away from this anchor sample. In other words, a hard positive sample $\mathbf{x}_{pos}$ and a hard negative sample $\mathbf{x}_{neg}$ with respect to the anchor sample $\mathbf{x}_{anc}$ can each be determined based on:
\begin{equation}
    {\left\| {{f\left(\mathbf{x}_{anc}\right)} - {f\left(\mathbf{x}_{pos}\right)}} \right\|_2^2} \geq \mathrm{t}_{pos}, \label{pos}
\end{equation}
\begin{equation}
    {\left\| {{f\left(\mathbf{x}_{anc}\right)} - {f\left(\mathbf{x}_{neg}\right)}} \right\|_2^2} \leq \mathrm{t}_{neg},\label{neg}
\end{equation}
where $\mathrm{t}_{pos}$ and $\mathrm{t}_{neg}$ respectively denote the thresholds that determine whether a sample is hard positive or hard negative. Since the hardest samples are likely to be outliers which can degrade the training process of the triplet network, we empirically set the $\mathrm{t}_{pos}$ at 75 percentile of the $L2$ distance and $\mathrm{t}_{neg}$ at 25 percentile of the $L2$ distance. The final set of $\mathbf{x}_{pos}$ and $\mathbf{x}_{neg}$ samples is randomly selected based on equations~\eqref{pos} and~\eqref{neg}.

@@

\subsubsection{Performance evaluation:}
The proposed system is evaluated under two cases: i) case I on comparison between the proposed system and other handcraft or deep learning-based methods based on the verification accuracy, and case II on detailed comparison between the proposed system and the deep learning-based methods using the receiver operating characteristic (ROC) curve and training loss curve. For both cases, Wi-Fi signal measurements based on a single-orientation and 4-orientations of the user are recorded. For Case I, existing handcraft methods such as the least squares error estimation (LSE)~\cite{duda2012pattern}, the principal components analysis (PCA) [22] with LSE, the support vector machine (SVM)~\cite{vapnik2013nature} with different kernel functions, the total error rate minimization which adopted the reduced multivariate polynomial model as basis function (TER-RM2) \cite{toh2003fingerprint,toh2008between}, the deep learning-based Siamese network \cite{koch2015siamese}, and the baseline triplet network \cite{hoffer2015deep} are included for performance benchmarking. The Siamese network and the baseline triplet network utilized the same ConvNet structures with the proposed system and only differed in their input data style and loss function.
